Together, Individuals Make a Difference

### 2025-01-27 论文/PAPER.tex
1

\begin{itemize}
	\setlength{\parsep}{0ex} %段落间距
	\setlength{\topsep}{2ex} %列表到上下文的垂直距离
	\setlength{\itemsep}{1ex} %条目间距
	\item Develop a model for each country’s medal count, which should at least include the number of gold medals and the total medal count, and estimate the uncertainty/precision of the model and evaluate its performance.
		\begin{itemize}
		\item[1)]
Using the model, predict the medal standings for the 2028 Los Angeles Summer Olympics (including prediction intervals for all outcomes). Based on the model’s predictions, identify which countries are most likely to improve their performance and which countries are expected to perform worse than in the 2024.
	\end{itemize}
	\begin{itemize}
		\item[2)]
		The developed model should include countries that have not yet won any medals, while also predicting the number of countries likely to win their first medal in the upcoming Olympics and provide odds for this estimate.
	\end{itemize}
		\begin{itemize}
		\item[3)]
The model should also consider the number and types of events in each specific Olympic Games, exploring the relationship between the events and the number of medals won by each country, as well as the impact of the events selected by the country on the results.
	\end{itemize}

	\item In contrast to athletes, who must change their citizenship in order to represent different countries in competitions, coaches do not need to change their citizenship to coach, which makes it easier for them to move to other countries. As a result, the great coach effect may arise. For example, Lang Ping has coached the volleyball teams of both China and the U.S., leading both to championships, while Bela Karolyi coached the Romanian and U.S. women’s gymnastics teams, achieving great success. The task is to identify evidence in the data that may suggest changes driven by the "great coach" effect and estimate its impact on the medal count. Select 3 countries, determine the events in which they should consider investing in "great" coaches, and estimate the impact of such investments
	\item Explain the unique insights regarding Olympic medal counts that are included in the developed model and how these insights can provide valuable information to the National Olympic Committees of various countries.

@@

The total medal count of a country more accurately reflects the long-term investments and developmental outcomes in sports, rather than short-term results that can be drastically altered. Furthermore, external factors that influence medal counts, such as the organization of international competitions, rule changes, and judging standards, tend to remain relatively stable in the short term.

@@

	\begin{longtable}{cl}
		\caption{Notations}\label{tb:notation} \\
		\toprule
		\multicolumn{1}{m{3cm}}{\centering Symbol}
		&\multicolumn{1}{m{8cm}}{\centering Definition}\\
		\midrule
		\endfirsthead

@@

To demonstrate the differences in the levels of countries in previous Olympic Games (i.e. national level), we employed the KMeans++ algorithm. Based on the number of gold, silver, and bronze medals, as well as the total medal count of each country in past Olympics, we categorized the countries into five levels. These 5 levels, L1, L2, L3, L4, and L5, form a partition of the sample set X (the total number of gold, silver, and bronze medals won by countries in previous Olympic Games).
\begin{equation}
G_{i} \cap G_{j}=\emptyset, \bigcup_{i=1}^{5} G_{i}=X
\end{equation}
After the classification, the distribution of countries across the 5 levels is shown in the figures:
\begin{figure}[H]
	\centering
	\includegraphics[width=12cm]{img/Level1.png}
	\caption{Scatter plot of national level classification (based on Kmeans++clustering algorithm)}
	\label{fig:aa}
\end{figure}

@@

In addition, we created 10 new features based on the original variables, which were utilized during the modeling process with the LightGBM model.

@@

\subsection{Data Mining}
\subsubsection{Athlete Service Status}
To predict future scenarios, we analyzed the trends in the number of participants for each country, event and year, and applied \textbf{Linear Regression} for fitting.

@@

From the distribution chart, it can be seen that nearly \( 80\% \)of athletes have participated in 1 or 2 matches, with the average number of serve times being 1.94, meaning each athlete participates in approximately 2 matches. Based on this assumption, we consider that each participating athlete will compete in 2 matches on average. Additionally, around \( 80\% \) of athletes have participated in only 1 Olympic Games, while nearly \( 20\% \) have participated in 2. Therefore, we can assume that \( 20\% \) of the veteran athletes will continue to compete in the next Olympic Games, providing a basis for determining which athletes will continue to participate.

To predict the Participant List, we employed Monte Carlo simulation, which consisted of two parts: the first part involved selecting \( 20\% \)of veteran athletes to continue competing (\( 20\% \) conforming to the original data distribution), while the second part generated new random athletes based on historical athlete data.

@@

For a given country in a specific event, the average number of medals won by all athletes serves as the standard number of medals for a representative athlete. The medal expectations were categorized into 3 scenarios: optimistic, moderate, and pessimistic. The optimistic expectation was calculated by averaging the top \( 50\% \) of the team, the moderate expectation by averaging all team members, and the pessimistic expectation by averaging the bottom \( 50\% \).

@@

By analyzing the performance of various countries in different events across past Olympic Games, we extracted a feature variable $\gamma$(Event Adeptness Coefficient), to represent the advantages of these countries in certain events.

@@

\subsubsection{Changes in Medal Count}
Since the IOC makes adjustments to events each year, the total number of medals awarded in each event does not remain exactly the same each year. To address this, we applied \textbf Linear Regression to fit the number of gold, silver, and bronze medals awarded in each event across past Olympic Games and predicted the medal distribution for the upcoming Olympic Games, providing a reference for subsequent operations.

\section{Task1:Medal Prediction Model Based on LightGBM}
\subsection{Medal Standings}
The medal table is influenced by various factors. To predict the standings for the next Olympics, we considered key factors identified during data preprocessing, such as nation level, participating athletes, country-specific event strengths, and adjustments to medal counts.

@@

And the steps of our LightGBM model are as below:
\begin{figure}[H]
	\centering
	\includegraphics[width=11cm]{img/xiao.png}
	\caption{LightGBM Model flow Chart}
	\label{fig:aa}
\end{figure}
The prediction of the Olympic medal table is essentially a multi-task learning problem, with the goal of predicting the number of gold, silver, and bronze medals each country is likely to win in different events. To accomplish this task, we need to input a large and diverse set of data while outputting predictions for multiple medal categories. These features are highly suitable for the LightGBM model.

Building on GBDT, LightGBM accelerates the computation using histograms, optimizing training efficiency and enabling the fast processing of high-dimensional, sparse, large-scale data. Compared to traditional logistic regression, LightGBM can capture nonlinear relationships through tree splitting, offering a performance advantage in accurately handling complex data and nonlinear relationships, significantly improving accuracy. Additionally, it can automatically handle missing values, providing great convenience for data preprocessing.

@@

\begin{equation}
	\mathcal{L} = \sum_{i=1}^{n} L(y_i, \hat{y}_i^{(m-1)} + \eta f_m(x_i)) + \Omega(f_m)
\end{equation}
The optimized loss function based on GBDT is represented as squared error in the medal prediction regression task.

@@

We input the previously obtained data into LightGBM for iterative model training. After training, we input the factors affecting the predictions for the next edition into the model, ultimately obtaining the predicted medal table for the 2028 Summer Olympics in Los Angeles, USA, are shown in Figure \ref{fig:USA}.
%奖牌榜预测图
\begin{figure}[H]
	\centering
	\includegraphics[width=11cm]{img/Predict.png}
	\caption{Medal Count Prediction}
	\label{fig:aa}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=12cm]{img/OP1.png}
	\caption{USA Medal count prediction}
	\label{fig:USA}
\end{figure}

According to the model's predictions, the United States is expected to see a significant increase in medal count compared to the previous year, while France will experience a notable decrease, which is related to the home-field advantage of the host country. Additionally, due to changes in the number of medals for certain events, the medal counts for Australia, Japan, and the Netherlands are expected to rise, while Italy and South Korea will see a decline. China's medal count is projected to remain relatively stable.

To more accurately simulate the actual number of medals won, we simplify each individual into 3 parameters: the probabilities of winning gold, silver, and bronze medals. For a specific country and event, the average number of medals won by all team members represents the probability that a standard "virtual athlete" should have.

@@

@@

Therefore, for countries that have never won a medal before, we calculate their first medal index $\xi$.

@@

The number of competitions is counted from the original dataset, and the expected increase in the number of medals for the next edition is predicted through a Linear Regression model based on historical data.

@@

In order to visualize the accuracy of a country's first medal prediction, we normalized the first medal index using the Z-score and then multiplied it by 0.8 to obtain the first medal coefficient $\kappa$.

@@

Here,$\overline{\xi}$ refers to the mean of the original data, and$\sigma$refers to the standard deviation of the original data.
The calculated results are shown in Figure \ref{fig:First Medal}:
%首奖可能性图
\begin{figure}[H]
	\centering
	\includegraphics[width=16cm]{img/First.png}
	\caption{The Probability of First Medal in 2028}
	\label{fig:First Medal}
\end{figure}

As shown in the figure, Sierra Leone (SLE), Antigua and Barbuda (ANT), and Papua New Guinea (PNG) are the three countries most likely to win their first medal in the next Olympic Games, with accuracies approximately 0.8, 0.77, and 0.74, respectively.

@@

@@

Here, $d_i$ is the rank difference of the two variables for the iii-th data point. 
Next, we compute the Spearman's Rank Correlation Coefficient ($R_s$).

@@

Through extensive data analysis, we identified 3 typical examples: table tennis in China, Athletics in Jamaica, and rowing in New Zealand, as shown in the figure.
%三个国家擅长项目的图
\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{.3\textwidth}
		\includegraphics[width=\textwidth]{img/Table Tennis.png}
		\caption{Heatmap of Correlation for Table Tennis in China}\label{subfig:1}
	\end{subfigure}
	\hfill 
	\begin{subfigure}[b]{.33\textwidth}
		\includegraphics[width=\textwidth]{img/Jamaica.png}
		\caption{ Heatmap of Correlation for Athletics in Jamaica}\label{subfig:2}
	\end{subfigure}
	\hfill 
	\begin{subfigure}[b]{.32\textwidth}
		\includegraphics[width=\textwidth]{img/New Zealand.png}
		\caption{Heatmap of Correlation for Rowing in New Zealand}\label{subfig:3}
	\end{subfigure}
	\caption{Three images}\label{fig:subfigures}
\end{figure}

We found that although China has a significant advantage in table tennis, the relationship between the development of the event and China's total medal count is relatively weak. This suggests that China's overall athletic adeptness is stable and less influenced by a single dominant event.

Additionally, Jamaica's total medal count shows a strong positive correlation with its performance in athletics, indicating the country's heavy reliance on athletics competitions. Since athletics is a well-established and stable event with minimal fluctuations in the number of events held each year, the relationship between the number of events and Jamaica's total medal count is relatively weak.

New Zealand also has a high dependency on rowing, with a substantial portion of its medals coming from this event. However, unlike Jamaica, rowing is not as established as athletics, and the number of competitions fluctuates from year to year. As a result, New Zealand's overall medal count is highly dependent on the rise and fall of rowing's prominence.
\subsection{Model Performances}

We calculated four evaluation metrics: Mean Squared Error (MSE), Root Mean Squared Error (RMSE), Mean Absolute Error (MAE), and the Coefficient of Determination (\(\mathbf{R}^2\)).

@@

\section{Task2:Great Coach Effect}
\subsection{Verification of the Great Coach Effect}
The Great Coach effect can cause significant differences in the performance of relevant countries in corresponding events. To collect evidence of changes caused by this effect, we use a Bayesian Changepoint Detection (BEAST) model.

@@

Using Bayes' theorem, we can compute the posterior probability for each possible change point $c_k$:

@@

\subsection{Countries Recommended for Hiring Coaches}
After confirming the significant impact of the great coach on the medal count, we began to analyze which events a country should consider investing in "excellent coaches." For this, we use the CUSUM algorithm.

@@

Firstly, the gender ratio in the Olympics has become increasingly balanced, with the male-to-female ratio nearly equal, shifting from a male-dominated ratio to approximately 1:1. The number of female athletes has steadily increased over time, particularly in the late 20th and early 21st centuries, with a more pronounced growth trend.

@@

In analyzing the national strengths in specific events, we found that certain countries have a dominant advantage, even to the point of forming a dominance, with their share of the total medal count reaching as high as \( 30\% \), \( 50\% \), or even over \( 80\% \).

\begin{figure}[H]
	\centering
	\includegraphics[width=16cm]{img/Monopolized Sports.png}
	\caption{Adeptness of Countries in Specitific Project}
	\label{fig:aa}
\end{figure}

@@

Based on the previously confirmed great coach effect, if other countries wish to enhance their capabilities in these events and win more medals, they may consider hiring coaches from these dominant countries.

@@

Sensitivity Analysis is used to evaluate the sensitivity of model outputs to changes in input features. Through this analysis, we can quantify the dependency of final prediction results prediction outcomes on specific data features, such as the adeptness of a particular country in a specific event or the previous award probability of an athlete.

LightGBM comes with a feature importance function that identifies the features with the greatest impact on the model’s predictions. By using plot\_importance() function, we can visualize the importance of each feature.

@@

Furthermore, we can use Shapley value analysis to assess the model's sensitivity. SHAP (SHapley Additive exPlanations) is a game theory-based model explanation method used to interpret machine learning model predictions. It assigns each feature a contribution to the model's prediction, helping us understand the importance and influence of each feature in the given prediction result.The formula for Shapley value is as follows:

@@

\(\phi_i(f)\) is the Shapley value of feature \(x_i\).\(S\) is a subset of features, and \(S \subseteq N \setminus \{i\}\) indicates that subset \(S\) contains all features except for \(x_i\).\(f(S)\) is the model's prediction on the feature subset \(S\).
\(f(S \cup \{i\})\) is the model's prediction on the subset \(S\) and feature \(x_i\).
\(|S|!\) and \((|N|-|S|-1)!\) are the combinatorial coefficients of the Shapley value, used to weigh the importance of each feature in different feature subsets.
\(|N|!\) is the total number of permutations of all features in the feature set \(N\), ensuring the average effect across all permutations.

@@

 we applied data preprocessing and feature engineering, using the \textbf{KMeans++} clustering algorithm to rank countries based on various features, including proficiency in different sports, host nation advantages, and athlete performance levels. The \textbf{LightGBM} regression model predicted the medal count for each country in the 2028 Olympics. Our analysis indicated that the United States is likely to progress, while France may decline, largely due to the home-field advantage. Countries such as Australia and Japan are expected to see an increase, while Italy and South Korea may experience a decrease. China's medal count is projected to remain stable. For nations without prior medals, we identified \textbf{Sierra Leone}, \textbf{Antigua and Barbuda}, and \textbf{Papua New Guinea} as those with the highest probability of winning their first medal in 2028. Correlation analysis revealed that countries with strong overall performance, like China, show minimal dependence on specific events, while smaller countries are more reliant on key events for their total medal count.

@@

 we explored the great coach effect using the \textbf{Bayesian Changepoint Detection} model. We found a strong correlation between coaching tenures and significant changes in a country’s sports performance. Additionally, the \textbf{CUSUM} algorithm helped identify weak events for several countries, such as France’s decline in sailing, Sweden’s waning dominance in athletics and shooting, and Germany’s emerging strength in swimming, which later showed signs of decline. Based on these findings, we recommend hiring specialized coaches to help maintain or regain competitive advantages in these events.
	\item \textbf{Task3}

 Using the models and data from Task 1 and Task 2, we performed a \textbf{detailed analysis} of several key trends in the Olympics. We found that the athlete \textbf{gender ratio} has become more balanced over time, reflecting increasing gender equality. Moreover, certain countries dominate \textbf{niche events}, such as the UK in cricket and Canada in lacrosse, which could offer opportunities for Olympic committees to enhance their medal prospects by investing in these specialized events. Lastly, our analysis of medal distribution changes over time revealed significant anomalies in 1904, 1980, 1984, and 1992, reflecting the Olympics' evolution and \textbf{geopolitical shifts}, particularly during the Cold War. These anomalies provide insights into broader global political and economic changes.
\end{itemize}

>>>>

\begin{itemize}
	\setlength{\parsep}{0ex} %段落间距
	\setlength{\topsep}{2ex} %列表到上下文的垂直距离
	\setlength{\itemsep}{1ex} %条目间距
	\item Develop a model for each country’s medal count, which should at least include the number of gold medals and the total medal count, and estimate the uncertainty/precision of the model and evaluate its performance.
		\begin{itemize}
		\item[1)]
Using the model, predict the medal table for the 2028 Los Angeles Summer Olympics (including prediction intervals for all outcomes). Based on the model’s predictions, identify which countries are most likely to improve their performance and which countries are expected to perform worse than in the 2024.
	\end{itemize}
	\begin{itemize}
		\item[2)]
		The developed model should include countries that have not yet won any medals, while also predicting the number of countries likely to win their first medal in the upcoming Olympics and provide odds for this estimate.
	\end{itemize}
		\begin{itemize}
		\item[3)]
The model should also consider the number and types of events in each specific Olympic Games, exploring the relationship between the events and the number of medals won by each country, as well as the impact of the events selected by the country on the results.
	\end{itemize}

	\item In contrast to athletes, who must change their citizenship in order to represent different countries in competitions, coaches do not need to change their citizenship to coach, which makes it easier for them to move to other countries. As a result, the great coach effect may arise. For example, Lang Ping has coached the volleyball teams of both China and the U.S., leading both to championships, while Bela Karolyi coached the Romanian and U.S. women’s gymnastics teams, achieving great success. The task is to identify evidence in the data that may suggest changes driven by the great coach effect and estimate its impact on the medal count. Select 3 countries, determine the events in which they should consider investing in "great" coaches, and estimate the impact of such investments
	\item Explain the unique insights regarding Olympic medal counts that are included in the developed model and how these insights can provide valuable information to the National Olympic Committees of various countries.

@@

The total medal count of a country more accurately reflects the long-term investments and developmental outcomes in sports, rather than short-term results that can be drastically altered. Furthermore, \textbf{external factors} that influence medal counts, such as the organization of international competitions, rule changes, and judging standards, tend to remain relatively stable in the short term.

@@

	\begin{longtable}{cl}
		\caption{Notations}\label{tb:notation} \\
		\toprule
		\multicolumn{1}{m{4cm}}{\centering Symbol}
		&\multicolumn{1}{m{13cm}}{\centering Definition}\\
		\midrule
		\endfirsthead

@@

To demonstrate the differences in the levels of countries in previous Olympic Games (i.e. national level), we employed the \textbf{KMeans++} algorithm. Based on the number of gold, silver, and bronze medals, as well as the total medal count of each country in past Olympics, we categorized the countries into five levels. These 5 levels, L1, L2, L3, L4, and L5, form a partition of the sample set X (the total number of gold, silver, and bronze medals won by countries in previous Olympic Games).
\begin{equation}
G_{i} \cap G_{j}=\emptyset, \bigcup_{i=1}^{5} G_{i}=X
\end{equation}
After the classification, the distribution of countries across the 5 levels is shown in the figures:
\begin{figure}[H]
	\centering
	\includegraphics[width=12cm]{img/Level1.png}
	\caption{Scatter plot of national level classification (based on KMeans++ clustering algorithm)}
	\label{fig:aa}
\end{figure}

@@

In addition, we created 10 new features based on the original variables, which were utilized during the modeling process with the \textbf{LightGBM} model.

@@

\subsection{Data Mining}
\subsubsection{Participant List Prediction}
To predict future scenarios, we analyzed the trends in the number of participants for each country, event and year, and applied \textbf{Linear Regression} for fitting.

@@

From the distribution chart, it can be seen that nearly \( 80\% \) of athletes have participated in 1 or 2 matches, with the average number of serve times being 1.94, meaning each athlete participates in approximately 2 matches. Based on this assumption, we consider that each participating athlete will compete in \textbf{2 matches on average}. Additionally, around \( 80\% \) of athletes have participated in only 1 Olympic Games, while nearly \( 20\% \) have participated in 2. Therefore, we can assume that \( 20\% \) of the veteran athletes will continue to compete in the next Olympic Games, providing a basis for determining which athletes will continue to participate.

To predict the Participant List, we employed Monte Carlo simulation, which consisted of two parts: the first part involved selecting \( 20\% \) of veteran athletes to continue competing (\( 20\% \) conforming to the original data distribution), while the second part generated new random athletes based on historical athlete data.

@@

For a given country in a specific event, the average number of medals won by all athletes serves as the standard number of medals for a representative athlete. The medal expectations were categorized into 3 scenarios: \textbf{optimistic}, \textbf{moderate}, and \textbf{pessimistic}. The optimistic expectation was calculated by averaging the top \( 50\% \) of the team, the moderate expectation by averaging all team members, and the pessimistic expectation by averaging the bottom \( 50\% \).

@@

By analyzing the performance of various countries in different events across past Olympic Games, we extracted a feature variable $\gamma$ (\textbf{Event Adeptness Coefficient}), to represent the advantages of these countries in certain events.

@@

\subsubsection{Changes in Medal Count}
Since the IOC makes adjustments to events each year, the total number of medals awarded in each event does not remain exactly the same each year. To address this, we applied\textbf{ Linear Regression} to fit the number of gold, silver, and bronze medals awarded in each event across past Olympic Games and predicted the medal distribution for the upcoming Olympic Games, providing a reference for subsequent operations.

\section{Task1:Medal Prediction Model Based on LightGBM}
\subsection{Medal Table}
The medal table is influenced by various factors. To predict the standings for the next Olympics, we considered key factors identified during data preprocessing, such as nation level, participating athletes, country-specific event strengths, and adjustments to medal counts.

@@

And the steps of our \textbf{LightGBM} model are as below:
\begin{figure}[H]
	\centering
	\includegraphics[width=11cm]{img/xiao.png}
	\caption{LightGBM Model Flow Chart}
	\label{fig:aa}
\end{figure}
The prediction of the Olympic medal table is essentially a multi-task learning problem, with the goal of predicting the number of gold, silver, and bronze medals each country is likely to win in different events. To accomplish this task, we need to input a large and diverse set of data while outputting predictions for multiple medal categories. These features are highly suitable for the LightGBM model.

Building on \textbf{GBDT}, LightGBM accelerates the computation using histograms, optimizing training efficiency and enabling the fast processing of high-dimensional, sparse, large-scale data. Compared to traditional logistic regression, LightGBM can capture \textbf{nonlinear relationships} through tree splitting, offering a performance advantage in accurately handling complex data and nonlinear relationships, significantly improving accuracy. Additionally, it can automatically handle missing values, providing great convenience for data preprocessing.

@@

\begin{equation}
	\mathcal{L} = \sum_{i=1}^{n} L(y_i, \hat{y}_i^{(m-1)} + \eta f_m(x_i)) + \Omega(f_m)
\end{equation}

The optimized loss function based on GBDT is represented as \textbf{squared error} in the medal prediction regression task.

@@

We input the previously obtained data into LightGBM for iterative model training. After training, we input the factors affecting the predictions for the next edition into the model, ultimately obtaining the predicted medal table for the 2028 Summer Olympics in Los Angeles, USA, are shown in Figure \ref{fig:aa}.
%奖牌榜预测图
\begin{figure}[H]
	\centering
	\includegraphics[width=11cm]{img/Predict.png}
	\caption{Medal Count Prediction}
	\label{fig:aa}
\end{figure}

According to the model's predictions, the United States is expected to see a significant increase in medal count compared to the previous year, while France will experience a notable decrease, which is related to the \textbf{home-field advantage} of the host country. Additionally, due to changes in the number of medals for certain events, the medal counts for Australia, Japan, and the Netherlands are expected to rise, while Italy and South Korea will see a decline. China's medal count is projected to remain relatively stable.

To more accurately simulate the actual number of medals won, we simplify each individual into 3 parameters: \textbf{the probabilities of winning gold, silver, and bronze medals}. For a specific country and event, the average number of medals won by all team members represents the probability that a standard "virtual athlete" should have.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=12cm]{img/OP1.png}
	\caption{USA Medal Count Prediction}
	\label{fig:USA}
\end{figure}

@@

Therefore, for countries that have never won a medal before, we calculate their \textbf{first medal index} $\xi$.

@@

The number of competitions is counted from the original dataset, and the expected increase in the number of medals for the next edition is predicted through a \textbf{Linear Regression model} based on historical data.

@@

In order to visualize the accuracy of a country's first medal prediction, we normalized the first medal index using the \textbf{Z-score} and then multiplied it by 0.8 to obtain the first medal coefficient $\kappa$.

@@

Here,$\overline{\xi}$ refers to the mean of the original data, and $\sigma$ refers to the standard deviation of the original data.
The calculated results are shown in Figure \ref{fig:First Medal}:
%首奖可能性图
\begin{figure}[H]
	\centering
	\includegraphics[width=16cm]{img/First.png}
	\caption{The Probability of First Medal in 2028}
	\label{fig:First Medal}
\end{figure}

As shown in the figure, \textbf{Sierra Leone (SLE)}, \textbf{Antigua and Barbuda (ANT)}, and \textbf{Papua New Guinea (PNG)} are the three countries most likely to win their first medal in the next Olympic Games, with accuracies approximately 0.8, 0.77, and 0.74, respectively.

@@

Here, $d_i$ is the rank difference of the two variables for the i-th data point. 
Next, we compute the \textbf{Spearman's Rank Correlation Coefficient} ($R_s$).

@@

@@

Through extensive data analysis, we identified 3 typical examples: table tennis in China, Athletics in Jamaica, and rowing in New Zealand, as shown in the figure.
%三个国家擅长项目的图
\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{.3\textwidth}
		\includegraphics[width=\textwidth]{img/Table Tennis.png}
		\caption{Heatmap of Correlation for Table Tennis in China}\label{subfig:1}
	\end{subfigure}
	\hfill 
	\begin{subfigure}[b]{.33\textwidth}
		\includegraphics[width=\textwidth]{img/Jamaica.png}
		\caption{ Heatmap of Correlation for Athletics in Jamaica}\label{subfig:2}
	\end{subfigure}
	\hfill 
	\begin{subfigure}[b]{.32\textwidth}
		\includegraphics[width=\textwidth]{img/New Zealand.png}
		\caption{Heatmap of Correlation for Rowing in New Zealand}\label{subfig:3}
	\end{subfigure}
	\caption{Heatmap of Correlation}\label{fig:subfigures}
\end{figure}

We found that although China has a significant advantage in table tennis, the relationship between the development of the event and China's total medal count is relatively weak. This suggests that China's overall athletic adeptness is \textbf{stable} and less influenced by a single dominant event.

Additionally, Jamaica's total medal count shows a strong positive correlation with its performance in athletics, indicating the country's heavy reliance on athletics competitions. Since athletics is a well-established and stable event with minimal fluctuations in the number of events held each year, the relationship between the number of events and Jamaica's total medal count is \textbf{relatively weak}.

New Zealand also has a high dependency on rowing, with a substantial portion of its medals coming from this event. However, unlike Jamaica, rowing is not as established as athletics, and the number of competitions fluctuates from year to year. As a result, New Zealand's overall medal count is highly \textbf{dependent} on the rise and fall of rowing's prominence.
\subsection{Model Performances}

We calculated four evaluation metrics: Mean Squared Error (\textbf{MSE}), Root Mean Squared Error (\textbf{RMSE}), Mean Absolute Error (\textbf{MAE}), and the Coefficient of Determination (\(\mathbf{R}^2\)).

@@

\section{Task2:Great Coach Effect}
\subsection{Verification of the Great Coach Effect}
The Great Coach effect can cause significant differences in the performance of relevant countries in corresponding events. To collect evidence of changes caused by this effect, we use a \textbf{Bayesian Changepoint Detection (BEAST)} model.

@@

Using Bayes' theorem, we can compute the \textbf{posterior probability} for each possible change point $c_k$:

@@

\subsection{Countries Recommended for Hiring Coaches}
After confirming the significant impact of the great coach on the medal count, we began to analyze which events a country should consider investing in great coaches. For this, we use the \textbf{CUSUM} algorithm.

@@

Firstly, the \textbf{gender ratio} in the Olympics has become increasingly \textbf{balanced}, with the male-to-female ratio nearly equal, shifting from a male-dominated ratio to approximately 1:1. The number of female athletes has steadily increased over time, particularly in the late 20th and early 21st centuries, with a more pronounced growth trend.

@@

In analyzing the national strengths in specific events, we found that certain countries have a \textbf{dominant advantage}, even to the point of forming a dominance, with their share of the total medal count reaching as high as \( 30\% \), \( 50\% \), or even over \textbf{\( 80\% \)}.

\begin{figure}[H]
	\centering
	\includegraphics[width=16cm]{img/Monopolized Sports.png}
	\caption{Adeptness of Countries in Specific Event}
	\label{fig:aa}
\end{figure}

@@

Based on the previously confirmed \textbf{great coach effect}, if other countries wish to enhance their capabilities in these events and win more medals, they may consider hiring coaches from these dominant countries.

@@

\textbf{Sensitivity Analysis} is used to evaluate the sensitivity of model outputs to changes in input features. Through this analysis, we can quantify the dependency of final prediction results prediction outcomes on specific data features, such as the adeptness of a particular country in a specific event or the previous award probability of an athlete.

LightGBM comes with a feature importance function that identifies the features with the greatest impact on the model’s predictions. By using \textbf{\textit{plot\_importance()}} function, we can visualize the importance of each feature.

@@

Furthermore, we can use Shapley value analysis to assess the model's sensitivity. \textbf{SHAP} (\textbf{SH}apley \textbf{A}dditive ex\textbf{P}lanations) is a game theory-based model explanation method used to interpret machine learning model predictions. It assigns each feature a contribution to the model's prediction, helping us understand the importance and influence of each feature in the given prediction result.The formula for Shapley value is as follows:

@@

\textbf{\(\phi_i(f)\)} is the Shapley value of feature \textbf{\(x_i\).\(S\)} is a subset of features, and \(S \subseteq N \setminus \{i\}\) indicates that subset \textbf{\(S\)} contains all features except for \textbf{\(x_i\).\(f(S)\)} is the model's prediction on the feature subset \textbf{\(S\)}.
\textbf{\(f(S \cup \{i\})\)} is the model's prediction on the subset \textbf{\(S\)} and feature \textbf{\(x_i\)}.

\textbf{\(|S|!\) }and \textbf{\((|N|-|S|-1)!\)} are the \textbf{combinatorial coefficients} of the Shapley value, used to weigh the importance of each feature in different feature subsets.
\textbf{\(|N|!\)} is the total number of permutations of all features in the feature set \textbf{\(N\)}, ensuring the average effect across all permutations.

@@

We applied data preprocessing and feature engineering, using the \textbf{KMeans++} clustering algorithm to rank countries based on various features, including proficiency in different sports, host nation advantages, and athlete performance levels.

The \textbf{LightGBM} regression model predicted the medal count for each country in the 2028 Olympics. Our analysis indicated that the United States is likely to progress, while France may decline, largely due to the home-field advantage. Countries such as Australia and Japan are expected to see an increase, while Italy and South Korea may experience a decrease.

China's medal count is projected to remain stable. For nations without prior medals, we identified \textbf{Sierra Leone}, \textbf{Antigua and Barbuda}, and \textbf{Papua New Guinea} as those with the highest probability of winning their first medal in 2028. Correlation analysis revealed that countries with strong overall performance, like China, show minimal dependence on specific events, while smaller countries are more reliant on key events for their total medal count.

@@

We explored the great coach effect using the \textbf{Bayesian Changepoint Detection} model. We found a strong correlation between coaching tenures and significant changes in a country’s sports performance.

Additionally, the \textbf{CUSUM} algorithm helped identify weak events for several countries, such as France’s decline in sailing, Sweden’s waning dominance in athletics and shooting, and Germany’s emerging strength in swimming, which later showed signs of decline. Based on these findings, we recommend hiring specialized coaches to help maintain or regain competitive advantages in these events.
	\item \textbf{Task3}

The models and data from Task 1 and Task 2, we performed a \textbf{detailed analysis} of several key trends in the Olympics. We found that the athlete \textbf{gender ratio} has become more balanced over time, reflecting increasing gender equality.

Moreover, certain countries dominate \textbf{niche events}, such as the UK in cricket and Canada in lacrosse, which could offer opportunities for Olympic committees to enhance their medal prospects by investing in these specialized events. Lastly, our analysis of medal distribution changes over time revealed significant anomalies in 1904, 1980, 1984, and 1992, reflecting the Olympics' evolution and \textbf{geopolitical shifts}, particularly during the Cold War. These anomalies provide insights into broader global political and economic changes.
\end{itemize}
